---
title: "Apology Assignment"
author: ""
embed-resources: true
format: html
---

# Task 1

## Step 1: Clean Data

In the raw data for this survey, there are originally 61 unique responses recorded (n = 61). Reviewing the data, one respondent (ID: R_bBJ75JWs3W8KSgF) was removed due to withdrawn consent and 12 respondents were removed due to vastly incomplete data and failed attention checks. In this data set there are two responses that are valuable but must be proceeded with caution. Respondent ID: R_3LfYCLoxkitDLMZ failed the attention check but the rest of the data is complete and adequate, there appears to have been a misunderstanding with how to answer the attention check, therefore this respondent has been left in to enrich the data. Respondent ID: R_1pRoB3KkAH3DjCD only has incomplete responses in the demographic section, yet has adequate and complete responses for the rest of the survey. After cleaning the data, the final set includes 48 unique responses (n = 48).

```{r}
library(tidyr)
library(ggplot2)
library(dplyr)

Cleaned_data <- read.csv("Cleaned_data.csv")
data_copy <- Cleaned_data
```

## Q1: Do People Care About a Return Apology?

```{r}
t.test(data_copy$feelings_youalone, data_copy$feelings_bothyoufirst, paired = TRUE)

mean_youalone <- round(mean(data_copy$feelings_youalone),2)
mean_bothyoufirst <- round(mean(data_copy$feelings_bothyoufirst),2)

cat("Average feelings score for 'No Return Apology':", mean_youalone, "\n")
cat("Average feelings score for 'Return Apology':", mean_bothyoufirst, "\n")

```

In the return apology scenario, individuals reported significantly more positive average feelings (7.44) compared to the scenario where individuals received no return apology (-18.4). The Paired T-test reveals that this difference is statistically significant as the p-value \< .05. Therefore, this data supports the alternative hypothesis that a return apology has an influence on feelings, suggesting people care about receiving a return apology.

### Return Apology Visualization

```         
    Paired t-test

data:  data_copy$feelings_youalone and data_copy$feelings_bothyoufirst
t = -12.789, df = 47, p-value < 2.2e-16
alternative hypothesis: true mean difference is not equal to 0
95 percent confidence interval:
 -29.89689 -21.76978
sample estimates:
mean difference 
      -25.83333 

Average feelings score for 'No Return Apology': -18.4 
Average feelings score for 'Return Apology': 7.44 
```

```{r}
return_apology <- gather(data_copy, variable, value, feelings_youalone, feelings_bothyoufirst)

ggplot(return_apology, aes(x = variable, y = value, fill = variable)) +
  geom_boxplot() +
  scale_x_discrete(labels = c("Return Apology", "No Return Apology")) +
  labs(title = "Comparison of Feelings with and Without Return Apology",
       x = "Condition", y = "Feelings",
       fill = "Condition") +
  scale_fill_manual(values = c("skyblue", "salmon"), 
                    labels = c("Return Apology", "No Return Apology")) +  
  theme_bw()
```

This visualization further depicts the magnitude of significant difference in feelings between receiving a return apology vs. receiving no return apology. It is apparent that respondents have a much more positive reaction to receiving a return apology, while the sentiment for no return apology is extremely negative.

## Q2: Does this vary as a function of Initiator Type?

### Magnitude of shift in feeling across Initiator Types

To measure the magnitude of shift in feelings, I transformed the data so all values can be positive in order to appropriately understand the direction and effect of the feelings between initiator types. It is important to proceed with caution when analyzing for initiator types as the sample size is a limitation to this analysis. Each initiator type is below a threshold of 30 respondents (always n=22, conditional n=16, never n=10), while this can provide useful insight, generalizing these results to the population may not provide full accuracy.

```{r}
return_apology$initiator_type <- as.factor(return_apology$initiator_type)
positive_data_copy <- data_copy %>%
  mutate(feelings_youalone_shifted = feelings_youalone + 30,
         feelings_bothyoufirst_shifted = feelings_bothyoufirst + 30)

positive_data_copy <- positive_data_copy %>%
  mutate(sentiment_difference = feelings_bothyoufirst_shifted - feelings_youalone_shifted)

anova <- aov(sentiment_difference ~ initiator_type, data = positive_data_copy)
summary(anova)
```

For this analysis, a one-way ANOVA test was used to compare means for multiple intitator types. The analysis depicts that p-value (0.677) \> alpha (0.05) for the shift in feelings across initiator types, suggesting that there is no significant difference in the magnitude of shift in feelings in the scenario of receiving a return apology compared to receiving no return apology among the initiator types.

### Difference in feelings for apology conditions across Initiator Types

```{r}
anova_no_return <- aov(feelings_youalone ~ initiator_type, data = data_copy)

cat("\nANOVA results for No Return Apology:\n ")
summary(anova_no_return)

cat("\nTukey Test results for No Return Apology:\n")
TukeyHSD(anova_no_return)

anova_yes_return <- aov(feelings_bothyoufirst ~ initiator_type, data = data_copy)

cat("\nANOVA results for Return Apology:\n")
summary(anova_yes_return)

cat("\nTukey Test results for Return Apology:\n")
print(TukeyHSD(anova_yes_return))


```

```{r}
data_long <- reshape2::melt(data_copy, id.vars = "initiator_type", 
                            measure.vars = c("feelings_youalone", "feelings_bothyoufirst"))

ggplot(data_long, aes(x = initiator_type, y = value, fill = variable)) +
  geom_boxplot(position = position_dodge(width = 0.8), color = "black") +
  labs(title = "Differences in Feelings for Apology Conditions Across Initiator Types",
       x = "Initiator Type", y = "Feelings") +
  scale_fill_manual(values = c("salmon", "skyblue"), 
                    labels = c("No Return Apology", "Return Apology")) +
  theme_minimal()
```

To measure the difference in feelings in each scenario (no return apology and return apology) between initiator types, a ANOVA test was used to compare the means. For the no return apology scenario, the p-value (.0305) \< alpha (.05) depicting that there is a significant difference between feelings across initiator types. A Tukey test was conducted to compare specific initiator types against each other. This analysis shows that in a scenario when no return apology is offered, there is a significant difference (p-value = 0.03) in feelings between those always willing to apologize firsts and those that would only be willing to apologize expecting a return apology.

For the return apology scenario, the p-value (.0399) \< alpha (.05) indicating that there is a significant difference between feelings across initiator types. The Tukey analysis depicts that there are no significant differences between types as the p-values are above the threshold of .05; however, the always-conditional comparison (p=.086) and always-never type comparison (p=.087) are close to the threshold. The average magnitude in these comparisons are quite high for the small sample size provided, this requires further investigation with a larger sample size.

## Q3: Is a Return Apology the same as Forgiving?

```{r}
t.test(data_copy$feelings_youaloneforgiven, data_copy$feelings_bothyoufirst, paired = TRUE)
mean_feelings_youaloneforgiven <- mean(data_copy$feelings_youaloneforgiven)

cat("Average feelings score for 'Offering Forgiveness':", mean_feelings_youaloneforgiven, "\n")
cat("Average feelings score for 'Return Apology':", mean_bothyoufirst, "\n")

```

In the return apology scenario, individuals on average reported significantly more positive feelings (7.44) compared to the scenario where individuals received forgiveness (-13.125). The Paired T-test reveals that this difference is statistically significant as the p-value \< .05. Therefore, this data supports the alternative hypothesis that receiving forgiveness is not viewed the same as a return apology.

# Task 2

## Q1: Average Feelings Visualization

```{r}
all_scenarios <- c("feelings_youalone","feelings_bothyoufirst","feelings_themalone","feelings_boththemfirst","feelings_neither","feelings_youaloneforgiven")

mean_scenarios <- colMeans(data_copy[, all_scenarios])
se_scenarios <- apply(data_copy[, all_scenarios], 2, sd) / sqrt(nrow(data_copy))

scenarios_df <- data.frame(
  scenario = names(mean_scenarios),
  mean = mean_scenarios,
  se = se_scenarios
)
scenarios_df$scenario <- as.factor(scenarios_df$scenario)

levels(scenarios_df$scenario) <- c("You Return", "They Return", "No Apology","Only Them", "Only You", "They Forgive")

ggplot(scenarios_df, aes(x = reorder(scenario, -mean), y = mean, fill = scenario)) +
  geom_bar(stat = "identity", position = "dodge") +
  geom_errorbar(aes(ymin = mean - se, ymax = mean + se), width = 0.2, position =   position_dodge(0.9)) +
  geom_text(aes(label = paste0("Mean: ",round(mean, 1), "\nSE: ± ", round(se, 1))),
            position = position_dodge(0.9), vjust = 2, size = 3) + 
  labs(title = "Average Feelings for Each Apology Scenario",
       x = "Apology Scenario",
       y = "Average Feelings") +
  theme_minimal()+ ylim(-30, 30)

```

The graph depicts that on average, return apologies elicit positive feelings, compared to more negative feelings evoked when only one party apologizes. This also illustrates how the other party taking initiative appears to reduce negative feelings; however, when the other party offers forgiveness, it elicits a similar magnitude of negative feelings as no parties apologizing.

## Q2: Difference in feelings across apology scenarios

```{r}
all_scenarios <- c("feelings_youalone", "feelings_bothyoufirst", "feelings_themalone",
                   "feelings_boththemfirst", "feelings_neither", "feelings_youaloneforgiven")

stacked_data <- stack(data_copy[, all_scenarios])

colnames(stacked_data) <- c("Feelings", "Scenario")

all_scenario_anova <- aov(Feelings ~ Scenario, data = stacked_data)

# Print ANOVA summary
summary(all_scenario_anova)
TukeyHSD(all_scenario_anova)

```

There are statistically significant differences in feelings across the various scenarios, supported by the ANOVA model's p-value \< . 001. The pairwise comparison indicates that reciprocal apologies are statistically significant and tend to generate more positive feelings when compared to scenarios without apologies or with one-sided apologies.

## Q3: Proportion of people choosing different options

### Preferred Outcome Visualization

```{r}
proportions <- data_copy %>%
  count(outcome_binary1) %>%
  mutate(proportion = n / sum(n))

ggplot(proportions, aes(x = "", y = proportion, fill = outcome_binary1)) +
  geom_bar(stat = "identity", width = 1) +
  coord_polar("y", start = 0) +
  labs(title = "Proportion of People's Apology Scenario Preferences",
       fill = "Preference") +
  theme_void() +
  theme(legend.position = "bottom") +
  scale_fill_manual(values = c("#Bca617", "#72dd83"), 
                    labels = c("They Return Apologize", "No One Apologize")) +
  geom_text(
    aes(label = paste0(round(proportion * 100, 1), "%")),
    position = position_stack(vjust = 0.5),
    size = 5
  )
```

### Significance Testing

```{r}
pref_table <- table(data_copy$outcome_binary1)

chi_squared_pref <- chisq.test(pref_table)

chi_squared_pref

```

Conducting a chi-squared goodness of fit test to evaluate distribution, the results show that since the p-value \< .001, we reject the null hypothesis, giving evidence to suggest that there is significant differences in the distribution of response across categories. The pie chart shows that 79.2% of respondents prefer a return apology after apologizing compared to no apology from either party, supporting the previous data assessing feelings across scenarios.

## Q4: NLP Exercise

The NLP code is in a separate HTML file, but I will keep the analysis here for simplicity. Using a text sentiment analysis, the text responses were split into 3 polarity labels: positive, negative, and neutral. Conducting an ANOVA test, there does not appear to be any significant impact on feelings for the different apology scenarios, suggesting that sentiment polarity may not be a strong indicator of variations in feelings. The graphs reinforce the notion that there is not a significant difference present; however, there are limitations due to the size of this data set.
